---
title: "Survey Geocode"
---

The purpose of this file is to geocode addresses provided by survey respondents
in the WUI survey. Addresses include the respondent's current address,
previous address, and work address. 

Load relevant libraries
```{r}
library(dplyr)
library(tidyr)
library(sf)
library(ggplot2)
library(tidygeocoder)
library(readr)
```

Import files for current, previous, and work addresses and delete rows that
are entirely NA
```{r}
curr_add <- read_csv("data/survey/current_addresses.csv")
curr_add <- curr_add[rowSums(is.na(curr_add)) < ncol(curr_add), ]

prev_add <- read_csv("data/survey/previous_addresses.csv")
prev_add <- prev_add[rowSums(is.na(prev_add)) < ncol(prev_add), ]

work_add <- read_csv("data/survey/work_addresses.csv")
work_add <- work_add[rowSums(is.na(work_add)) < ncol(work_add), ]
```

Our current address file has location information for where all respondents
currently live, and it has location information for where **some** respondents
previously lived, and where they currently work. The previous and work address
files have that address information for other respondents.

Each respondent has a global ID. Let's use this ID to merge the three address
datasets together so that we don't have to deal with three separate data frames.

```{r}
# Join work address and previous address
all_add <- left_join(curr_add, prev_add, by = "ParentGlobalID")

# Join work and previous addresses to work addresses
all_add <- left_join(all_add, work_add, by = "ParentGlobalID")
```

Previous and work addresses that start with "m_" came from the current address
data frame, and those start with "m2_" came from the previous or work address
sheet. Let's combine the "m_" and "m2_" addresses into single columns to work
with.

```{r}
all_add1 <- all_add %>%
  mutate(prev_adr = case_when(!is.na(m_prev_adr) & is.na(m2_prev_adr) ~ m_prev_adr,
                              is.na(m_prev_adr) & !is.na(m2_prev_adr) ~ m2_prev_adr,
                              m_prev_adr == m2_prev_adr ~ m_prev_adr,
                              !is.na(m_prev_adr) & !is.na(m2_prev_adr) & 
                                m_prev_adr != m2_prev_adr ~ "CONFLICT",
                              TRUE ~ NA),
         prev_cpn = case_when(!is.na(m_prev_cpn) & is.na(m2_prev_cpn) ~ m_prev_cpn,
                              is.na(m_prev_cpn) & !is.na(m2_prev_cpn) ~ m2_prev_cpn,
                              m_prev_cpn == m2_prev_cpn ~ m_prev_cpn,
                              !is.na(m_prev_cpn) & !is.na(m2_prev_cpn) & 
                                m_prev_cpn != m2_prev_cpn ~ "CONFLICT",
                              TRUE ~ NA),
         prev_zip = case_when(!is.na(m_prev_zip) & is.na(m2_prev_zip) ~ m_prev_zip,
                              is.na(m_prev_zip) & !is.na(m2_prev_zip) ~ m2_prev_zip,
                              m_prev_zip == m2_prev_zip ~ m_prev_zip,
                              !is.na(m_prev_zip) & !is.na(m2_prev_zip) & 
                                m_prev_zip != m2_prev_zip ~ 99999,
                              TRUE ~ NA),
         prev_state = case_when(!is.na(m_prev_state) & is.na(m2_prev_state) ~ m_prev_state,
                              is.na(m_prev_state) & !is.na(m2_prev_state) ~ m2_prev_state,
                              m_prev_state == m2_prev_state ~ m_prev_state,
                              !is.na(m_prev_state) & !is.na(m2_prev_state) & 
                                m_prev_state != m2_prev_state ~ "CONFLICT",
                              TRUE ~ NA),
         prev_cntry = case_when(!is.na(m_prev_cntry) & is.na(m2_prev_cntry) ~ m_prev_cntry,
                              is.na(m_prev_cntry) & !is.na(m2_prev_cntry) ~ m2_prev_cntry,
                              m_prev_cntry == m2_prev_cntry ~ m_prev_cntry,
                              !is.na(m_prev_cntry) & !is.na(m2_prev_cntry) & 
                                m_prev_cntry != m2_prev_cntry ~ "CONFLICT",
                              TRUE ~ NA),
         prev_lat = case_when(!is.na(m_prev_LAT) & is.na(m2_prev_LAT) ~ m_prev_LAT,
                            is.na(m_prev_LAT) & !is.na(m2_prev_LAT) ~ m2_prev_LAT,
                            m_prev_LAT == m2_prev_LAT ~ m_prev_LAT,
                           !is.na(m_prev_LAT) & !is.na(m2_prev_LAT) & 
                               m_prev_LAT != m2_prev_LAT &
                               (m_prev_LAT - m2_prev_LAT <= .01) #keep m2 if the difference is small
                                & (m_prev_LAT - m2_prev_LAT >= -.01) ~ m2_prev_LAT,
                            !is.na(m_prev_LAT) & !is.na(m2_prev_LAT) & 
                               m_prev_LAT != m2_prev_LAT &
                               (m_prev_LAT - m2_prev_LAT >= .01) #conflict if the difference is > .01
                                & (m_prev_LAT - m2_prev_LAT <= -.01) ~ 99999,
                            TRUE ~ NA),
         prev_long = case_when(!is.na(m_prev_LONG) & is.na(m2_prev_LONG) ~ m_prev_LONG,
                            is.na(m_prev_LONG) & !is.na(m2_prev_LONG) ~ m2_prev_LONG,
                            m_prev_LONG == m2_prev_LONG ~ m_prev_LONG,
                            !is.na(m_prev_LONG) & !is.na(m2_prev_LONG) & 
                               m_prev_LONG != m2_prev_LONG &
                               (m_prev_LONG - m2_prev_LONG <= .01) #keep m2 if the difference is small
                                & (m_prev_LONG - m2_prev_LONG >= -.01) ~ m2_prev_LONG,
                            !is.na(m_prev_LONG) & !is.na(m2_prev_LONG) & 
                               m_prev_LONG != m2_prev_LONG &
                               (m_prev_LONG - m2_prev_LONG >= .01) #conflict if the difference is > .01
                                & (m_prev_LONG - m2_prev_LONG <= -.01) ~ 99999,
                            TRUE ~ NA),
         work_adr = case_when(!is.na(m_w_adr) & is.na(m2_work_adr) ~ m_w_adr,
                              is.na(m_w_adr) & !is.na(m2_work_adr) ~ m2_work_adr,
                              m_w_adr == m2_work_adr ~ m_w_adr,
                              !is.na(m_w_adr) & !is.na(m2_work_adr) & 
                                m_w_adr != m2_work_adr ~ "CONFLICT",
                              TRUE ~ NA),
         work_cpn = case_when(!is.na(m_w_cpn) & is.na(m2_work_cpn) ~ m_w_cpn,
                              is.na(m_w_cpn) & !is.na(m2_work_cpn) ~ m2_work_cpn,
                              m_w_cpn == m2_work_cpn ~ m_w_cpn,
                              !is.na(m_w_cpn) & !is.na(m2_work_cpn) & 
                                m_w_cpn != m2_work_cpn ~ "CONFLICT",
                              TRUE ~ NA),
         work_zip = case_when(!is.na(m_w_zip) & is.na(m2_work_zip) ~ m_w_zip,
                              is.na(m_w_zip) & !is.na(m2_work_zip) ~ as.numeric(m2_work_zip),
                              m_w_zip == as.numeric(m2_work_zip) ~ m_w_zip,
                              !is.na(m_w_zip) & !is.na(m2_work_zip) & 
                                m_w_zip != as.numeric(m2_work_zip) ~ 99999,
                              TRUE ~ NA),
         work_state = case_when(!is.na(m_w_state) & is.na(m2_work_state) ~ m_w_state,
                              is.na(m_w_state) & !is.na(m2_work_state) ~ m2_work_state,
                              m_w_state == m2_work_state ~ m_w_state,
                              !is.na(m_w_state) & !is.na(m2_work_state) & 
                                m_w_state != m2_work_state ~ "CONFLICT",
                              TRUE ~ NA),
         work_cntry = m2_work_cntry, #No country value in the "m_w" set of vars
         work_lat = case_when(!is.na(m_w_LAT) & is.na(m2_work_LAT) ~ m_w_LAT,
                            is.na(m_w_LAT) & !is.na(m2_work_LAT) ~ m2_work_LAT,
                            m_w_LAT == m2_work_LAT ~ m_w_LAT,
                            !is.na(m_w_LAT) & !is.na(m2_work_LAT) & 
                               m_w_LAT != m2_work_LAT &
                               (m_w_LAT - m2_work_LAT <= .01) #keep m2 if the difference is small
                                & (m_w_LAT - m2_work_LAT >= -.01) ~ m2_work_LAT,
                           !is.na(m_w_LAT) & !is.na(m2_work_LAT) & 
                               m_w_LAT != m2_work_LAT &
                               (m_w_LAT - m2_work_LAT >= .01) #conflict if the difference is > .01
                                & (m_w_LAT - m2_work_LAT <= -.01) ~ 99999,
                            TRUE ~ NA),
         work_long = case_when(!is.na(m_w_LONG) & is.na(m2_work_LONG) ~ m_w_LONG,
                            is.na(m_w_LONG) & !is.na(m2_work_LONG) ~ m2_work_LONG,
                            m_w_LONG == m2_work_LONG ~ m_w_LONG,
                            !is.na(m_w_LONG) & !is.na(m2_work_LONG) & 
                               m_w_LONG != m2_work_LONG &
                               (m_w_LONG - m2_work_LONG <= .01) #keep m2 if the difference is small
                                & (m_w_LONG - m2_work_LONG >= -.01) ~ m2_work_LONG,
                           !is.na(m_w_LONG) & !is.na(m2_work_LONG) & 
                               m_w_LONG != m2_work_LONG &
                               (m_w_LONG - m2_work_LONG >= .01) #conflict if the difference is > .01
                                & (m_w_LONG - m2_work_LONG <= -.01) ~ 99999,
                            TRUE ~ NA),
         
         )

```


Before we geocode, we want an indicator variable for each address that says
whether or not an address is a partial address. This is important because when
presenting the locations of survey respondents, we want to know the degree of
spatial uncertainty we're working with. An address will be marked as partial if
it's missing street information AND it's 1) missing XY coordinates from the 
respondent dropping a pin in their location or 2) missing XY coordinates manually
filled in from the survey team, which generally occurs when a respondent puts in
a street name but not an address number. This will not account for addresses, 
however, that have a street name but no address number.
```{r}
all_add2 <- all_add1 %>%
  mutate(c_partial = case_when(is.na(m_CURR_ADR) & ((is.na(m_CURR_LAT) | 
                                                     is.na(m_CURR_LONG)) &
                                                    (is.na(p_CURR_LAT) |
                                                     is.na(p_CURR_LONG))
                                                    ) ~ 1,
                               TRUE ~ 0),
         p_partial = case_when(is.na(prev_adr) & ((is.na(prev_lat) | 
                                                     is.na(prev_long)) &
                                                    (is.na(p_prev_LAT) |
                                                     is.na(p_prev_LONG))
                                                    ) ~ 1,
                               TRUE ~ 0),
         w_partial = case_when(is.na(work_adr) & ((is.na(work_lat) | 
                                                     is.na(work_long)) &
                                                    (is.na(p_w_LAT) |
                                                     is.na(p_w_LONG))
                                                    ) ~ 1,
                               TRUE ~ 0)
  )
```

It appears that Open Street Map does a better job of geocoding partial addresses
when they're in a single line address format than it does with combining
components of an address together, so let's put all three addresses on a single
line.

```{r}
all_add3 <- all_add2 %>% 
  unite("curr_sla", c(m_CURR_ADR, m_CURR_CPN, m_CURR_STATE), 
        sep=", ", remove = FALSE, na.rm = TRUE) %>%
  unite("curr_slaz", c(curr_sla, m_CURR_ZIP), 
        sep=" ", remove = FALSE, na.rm = TRUE) %>%
  unite("prev_sla", c(prev_adr, prev_cpn, prev_state), 
        sep=", ", remove = FALSE, na.rm = TRUE) %>%
  unite("prev_slaz", c(prev_sla, prev_zip), 
        sep=" ", remove = FALSE, na.rm = TRUE) %>%
  unite("work_sla", c(work_adr, work_cpn, work_state), 
        sep=", ", remove = FALSE, na.rm = TRUE) %>%
  unite("work_slaz", c(work_sla, work_zip), 
        sep=" ", remove = FALSE, na.rm = TRUE)
```

Now, let's geocode the current addresses. We'll ask the census to geocode each
address, and for the addresses that the census can't geocode, we'll ask Open
Street Map.

```{r}
# Go back and try this with a single dataframe with piping
gc1 <- all_add3 %>%
  geocode_combine(
    queries = list(
      list(method = 'census'),
      list(method = 'osm')
    ),
    global_params = list(address = 'curr_slaz'),
    lat = "cgc_lat",
    lon = "cgc_lon") %>%
  rename(c_query = query)

gc2 <- gc1 %>%
  geocode_combine(
    queries = list(
      list(method = 'census'),
      list(method = 'osm')
    ),
    global_params = list(address = 'prev_slaz'),
    lat = "pgc_lat",
    lon = "pgc_lon") %>%
  rename(p_query = query)
  
gc3 <- gc2 %>%
  geocode_combine(
    queries = list(
      list(method = 'census'),
      list(method = 'osm')
    ),
    global_params = list(address = 'work_slaz'),
    lat = "wgc_lat",
    lon = "wgc_lon") %>%
  rename(w_query = query)
```

Let's save a copy of gc3 so that the survey team can review the addresses 
that didn't geocode.
```{r}
write_csv(gc3, "data/survey/geocode08-20-24.csv")
```

So the last thing to do here is create final coordinates columns for current,
previous, and work addresses. There are two options for the coordinates to put
in this column: the geocoded coordinates or, if the respondent told us their
location by dropping a pin in the map, the coordinates the respondent gave us.

This is the logic we want to follow:

1. If there are manually entered coordinates, use those. This accounts for
addresses that lack a street number but aren't considered partial.
2. If respondent provided a complete address, use the geocoded coordinates
(including if the respondent dropped a pin)
3. If the respondent did not provide a complete address but dropped a pin, use
the pin coordinates.
4. If the respondent provided a complete address and dropped a pin, but we couldn't
geocode the complete address, use the pin coordinates.
5. If the respondent did not provide a complete address and did not drop a pin,
but we were able to geocode the incomplete address, use the geocoded coordinates

```{r}
gc4 <- gc3 %>%
  mutate(cc_lat = case_when(!is.na(m_CURR_LAT) ~ m_CURR_LAT,
                            c_partial == 0 & !is.na(cgc_lat) ~ cgc_lat,
                            c_partial == 1 & !is.na(p_CURR_LAT) ~ p_CURR_LAT,
                            c_partial == 0 & is.na(cgc_lat) & !is.na(p_CURR_LAT) ~ p_CURR_LAT,
                            c_partial == 1 & is.na(p_CURR_LAT) & !is.na(cgc_lat) ~ cgc_lat,
                            TRUE ~ NA),
         cc_lon = case_when(!is.na(m_CURR_LONG) ~ m_CURR_LONG,
                            c_partial == 0 & !is.na(cgc_lon) ~ cgc_lon,
                            c_partial == 1 & !is.na(p_CURR_LONG) ~ p_CURR_LONG,
                            c_partial == 0 & is.na(cgc_lon) & !is.na(p_CURR_LONG) ~ p_CURR_LONG,
                            c_partial == 1 & is.na(p_CURR_LONG) & !is.na(cgc_lon) ~ cgc_lon,
                            TRUE ~ NA),
         cp_lat = case_when(!is.na(prev_lat) ~ prev_lat,
                            p_partial == 0 & !is.na(pgc_lat) ~ pgc_lat,
                            p_partial == 1 & !is.na(p_prev_LAT) ~ p_prev_LAT,
                            p_partial == 0 & is.na(pgc_lat) & !is.na(p_prev_LAT) ~ p_prev_LAT,
                            p_partial == 1 & is.na(p_prev_LAT) & !is.na(pgc_lat) ~ pgc_lat,
                            TRUE ~ NA),
         cp_lon = case_when(!is.na(prev_long) ~ prev_long,
                            p_partial == 0 & !is.na(pgc_lon) ~ pgc_lon,
                            p_partial == 1 & !is.na(p_prev_LONG) ~ p_prev_LONG,
                            p_partial == 0 & is.na(pgc_lon) & !is.na(p_prev_LONG) ~ p_prev_LONG,
                            p_partial == 1 & is.na(p_prev_LONG) & !is.na(pgc_lon) ~ pgc_lon,
                            TRUE ~ NA),
         cw_lat = case_when(!is.na(work_lat) ~ work_lat,
                            w_partial == 0 & !is.na(wgc_lat) ~ wgc_lat,
                            w_partial == 1 & !is.na(p_w_LAT) ~ p_w_LAT,
                            w_partial == 0 & is.na(wgc_lat) & !is.na(p_w_LAT) ~ p_w_LAT,
                            w_partial == 1 & is.na(p_w_LAT) & !is.na(wgc_lat) ~ wgc_lat,
                            TRUE ~ NA),
         cw_lon = case_when(!is.na(work_long) ~ work_long,
                            w_partial == 0 & !is.na(wgc_lon) ~ wgc_lon,
                            w_partial == 1 & !is.na(p_w_LONG) ~ p_w_LONG,
                            w_partial == 0 & is.na(wgc_lon) & !is.na(p_w_LONG) ~ p_w_LONG,
                            w_partial == 1 & is.na(p_w_LONG) & !is.na(wgc_lon) ~ wgc_lon,
                            TRUE ~ NA),
  )
  
```

Now we just want to wittle down out dataframe to variables we need to integrate
with WUI data in a different .Rmd file
```{r}
gc5 <- gc4 %>%
  select(ParentGlobalID, c_partial, p_partial, w_partial, cc_lat, cc_lon, cp_lat, cp_lon,
         cw_lat, cw_lon) %>%
  rename(curr_lat = cc_lat,
         curr_lon = cc_lon,
         prev_lat = cp_lat,
         prev_lon = cp_lon,
         work_lat = cw_lat,
         work_lon = cw_lon)
```


Now let's save this as a CSV for later use.
```{r}
write_csv(gc5, "data/survey/survey_coords.csv")
```

