---
title: "Survey WUI integration"
---

Import relevant libraries
```{r}
library(dplyr)
library(sf)
library(ggplot2)
library(nngeo)
library(readr)
```

Read in survey address coordinates
```{r}
survey_xy <- read_csv("data/survey/survey_coords.csv")
```

Let's subset the data into three dataframes. We'll need to do this to project
the coordinates for current, previous, and work addresses. We'll merge the
matching observations back together at the end, ending up with three separate
SF objects (current addresses, current + previous addresses, and current 
+ work addresses).

```{r}
curr_xy <- survey_xy %>%
  select(ParentGlobalID, c_partial, curr_lat, curr_lon) 

#Get only complete cases because we can't project NA values
curr_xy <- curr_xy[complete.cases(curr_xy), ]

prev_xy <- survey_xy %>%
  select(ParentGlobalID, p_partial, prev_lat, prev_lon)

prev_xy <- prev_xy[complete.cases(prev_xy), ]

work_xy <- survey_xy %>%
  select(ParentGlobalID, w_partial, work_lat, work_lon)

work_xy <- work_xy[complete.cases(work_xy), ]
```

Convert coordinates df into sf objects
```{r}
c_points <- st_as_sf(curr_xy, coords = c('curr_lon', 'curr_lat'), crs = 4326)
p_points <- st_as_sf(prev_xy, coords = c('prev_lon', 'prev_lat'), crs = 4326)
w_points <- st_as_sf(work_xy, coords = c('work_lon', 'work_lat'), crs = 4326)
```


Import necessary polygon datasets
```{r}
counties <- st_read("data/shapes/4county_boundaries.shp")
places <- st_read("data/shapes/4county_places.shp")
wui_bg <- st_read("data/shapes/wui_bg_demos1.geojson")
wui_blks <- st_read("data/shapes/ca_wui_pro.geojson")
sc_usl <- st_read("data/shapes/santa_cruz_urban_services_line.geojson")
```

Project our survey points into the same coordinate system as our polygons
```{r}

c_points <- st_transform(c_points, st_crs(counties))
st_crs(c_points) == st_crs(counties)

p_points <- st_transform(p_points, st_crs(counties))
st_crs(p_points) == st_crs(counties)

w_points <- st_transform(w_points, st_crs(counties))
st_crs(w_points) == st_crs(counties)
```

## Join counties to points

Let's join points to counties. This will tell us which points are assigned to
which counties, but it won't tell us the distance between the points and 
counties, which is necessary for us to know how many survey respondents lived 
inside versus outside a county.

First, let's reduce the number of variables in our county object to GEOID and
county name
```{r}
counties1 <- counties %>%
  rename(county_fips = GEOID,
         county = NAME) %>%
  select(county_fips, county)
```

Join our counties to our points
```{r}
county_to_c_point <- st_join(c_points, counties1, join = st_nn, k=1)
county_to_p_point <- st_join(p_points, counties1, join = st_nn, k=1)
county_to_w_point <- st_join(w_points, counties1, join = st_nn, k=1)
```
Now let's get our distances
```{r}
c_cnty_n = st_nn(c_points, counties, k=1, returnDist = TRUE)
c_cnty_dist = sapply(c_cnty_n[[2]], "[", 1)

p_cnty_n = st_nn(p_points, counties, k=1, returnDist = TRUE)
p_cnty_dist = sapply(p_cnty_n[[2]], "[", 1)

w_cnty_n = st_nn(w_points, counties, k=1, returnDist = TRUE)
w_cnty_dist = sapply(w_cnty_n[[2]], "[", 1)

c_cnty_dist
```

Now lets incorporate our distances to our county data
```{r}
county_to_c_point$c_cnty_dist = c_cnty_dist
county_to_p_point$p_cnty_dist = p_cnty_dist
county_to_w_point$w_cnty_dist = w_cnty_dist
```

Now let's create a text column that will, if the respondent lives INSIDE a 
county, will return the name of that county, and if the respondent lives OUTSIDE
a county, will return "X km from Y county".

```{r}
county_to_c_point1 <- county_to_c_point %>%
  mutate(c_inside_cnty = case_when(c_cnty_dist == 0 ~ "Yes",
                                 c_cnty_dist > 0 ~ paste(round(c_cnty_dist/1000, digits = 1), "km from ", county, sep = ""),
                                 TRUE ~ "ERROR"))

county_to_p_point1 <- county_to_p_point %>%
  mutate(p_inside_cnty = case_when(p_cnty_dist == 0 ~ "Yes",
                                 p_cnty_dist > 0 ~ paste(round(p_cnty_dist/1000, digits = 1), "km from ", county, sep = ""),
                                 TRUE ~ "ERROR"))

county_to_w_point1 <- county_to_w_point %>%
  mutate(w_inside_cnty = case_when(w_cnty_dist == 0 ~ "Yes",
                                 w_cnty_dist > 0 ~ paste(round(w_cnty_dist/1000, digits = 1), "km from ", county, sep = ""),
                                 TRUE ~ "ERROR"))
```

## Join places to points

Let's shave down our places object to just place names and their FIPS codes
```{r}
places1 <- places %>%
  rename(pl_fips = GEOID,
         place = NAMELSAD) %>%
  select(pl_fips, place)
```

Let's join points to places. This will tell us which points are assigned to
which places, but it won't tell us the distance between the points and places,
which is necessary for us to know how many survey respondents lived inside 
versus outside a place.
```{r}
places_to_c_point <- st_join(county_to_c_point1, places1, join = st_nn, k=1)
places_to_p_point <- st_join(county_to_p_point1, places1, join = st_nn, k=1)
places_to_w_point <- st_join(county_to_w_point1, places1, join = st_nn, k=1)
```

Now let's get our place distances
```{r}
c_pl_n = st_nn(places_to_c_point, places1, k=1, returnDist = TRUE)
c_pl_dist = sapply(c_pl_n[[2]], "[", 1)


p_pl_n = st_nn(places_to_p_point, places1, k=1, returnDist = TRUE)
p_pl_dist = sapply(p_pl_n[[2]], "[", 1)


w_pl_n = st_nn(places_to_w_point, places1, k=1, returnDist = TRUE)
w_pl_dist = sapply(w_pl_n[[2]], "[", 1)

c_pl_dist
```

Now lets bind our distances to our place data
```{r}
places_to_c_point$c_pl_dist = c_pl_dist
places_to_p_point$p_pl_dist = p_pl_dist
places_to_w_point$w_pl_dist = w_pl_dist
```

Now let's create a text column that will, if the respondent lives INSIDE a place,
will return the name of that place, and if the respondent lives OUTSIDE a place,
will return "Unincorporated near place name".

```{r}
places_to_c_point1 <- places_to_c_point %>%
  mutate(c_pl_res = case_when(c_pl_dist == 0 ~ place,
                            c_pl_dist > 0 ~ paste("Unincorporated near ", 
                                                place, sep = ""),
                            TRUE ~ "ERROR"))

places_to_p_point1 <- places_to_p_point %>%
  mutate(p_pl_res = case_when(p_pl_dist == 0 ~ place,
                            p_pl_dist > 0 ~ paste("Unincorporated near ", 
                                                place, sep = ""),
                            TRUE ~ "ERROR"))

places_to_w_point1 <- places_to_w_point %>%
  mutate(pl_work = case_when(w_pl_dist == 0 ~ place,
                            w_pl_dist > 0 ~ paste("Unincorporated near ", 
                                                place, sep = ""),
                            TRUE ~ "ERROR"))

```

## Determine whether points are inside the Santa Cruz County urban services line

Convert urban service line CRS to same as our point data
```{r}
sc_usl <- st_transform(sc_usl, st_crs(places_to_c_point1))
st_crs(places_to_c_point1) == st_crs(sc_usl)
```

```{r}
usl_to_c_point <- places_to_c_point1
usl_to_p_point <- places_to_p_point1
usl_to_w_point <- places_to_w_point1

usl_to_c_point$sc_usl = case_when(st_intersects(usl_to_c_point, sc_usl, sparse = F) ~ 1,
                                  TRUE ~ 0)
usl_to_p_point$sc_usl = case_when(st_intersects(usl_to_p_point, sc_usl, sparse = F) ~ 1,
                                  TRUE ~ 0)
usl_to_w_point$sc_usl = case_when(st_intersects(usl_to_w_point, sc_usl, sparse = F) ~ 1,
                                  TRUE ~ 0)
```

## Join WUI status to points (Block)

Here, we want to determine whether a respondent lives, previously lives,
or works in the WUI (even outside our study area, but not necessarily outside
CA). To do this, let's create an sf object with five variables: whether the 
block was in the WUI in 2020, whether the block was intermix in 2020, whether 
the block was interface in 2020, whether the block was new WUI 
(extensive growth) in 2020, and block ID. We'll merge those variables into our 
datasets of surveyor locations.

```{r}
wui_status20 <- wui_blks %>%
  select(BLK20, wui20, intermix20, interfac20, wui_new20, high_den20, med_den20,
         low_den20)
```

Let's join points to blocks. This will tell us which points are assigned to
which blocks and their associated WUI status.
```{r}
places_to_c_point2 <- st_join(usl_to_c_point, wui_status20)
places_to_p_point2 <- st_join(usl_to_p_point, wui_status20)
places_to_w_point2 <- st_join(usl_to_w_point, wui_status20)
```

## Join growth WUI status to points (Block Group)

Here, we will use block group data to understand whether a survey respondent
lives in an area of intensifying WUI growth. We're using the block group because
it is difficult to trust housing unit growth data at the block level. However,
we have to establish thresholds: how do we decide what block groups are part of
the WUI? Do we use WUI area or WUI housing units as the criteria? What are the
thresholds we might use for each? And what is the threshold for housing
density growth that classifies a WUI block group as experiencing intensifying
growth?

For now, let's say WUI block groups are blocks where 50% or more of the housing
units are in the WUI. This is a more liberal definition than one based on area.
For example, this definition classifies more block groups as part of the WUI 
than one that defines the as block groups where 25% of more of the area is part
of the WUI. 

Let's define intensification along two criteria: housing density growth
and population density growth. We'll break those each up into two thresholds:
light growth and heavy growth. The respective thresholds for each will be five
percent and ten percent.

Let's create indicator variables to tell us:
1. Whether a block group is part of the WUI based on our 50% of housing units
criteria.
2. Whether a block group experienced at least 5% growth in housing units from
2010-2020 (light intensification).
3. Whether a block group experienced at least 10% growth in housing units from
2010-2020 (heavy intensification)
4. Whether a block group experienced at least 5% growth in population density
from 2010-2020 (light intensification).
5 Whether a block group experienced at least 10% growth in population density
from 2010-2020 (heavy intensification).

And then we will subset our data down to relevant columns to be joined to the
point data.
```{r}
wui_bg1 <- wui_bg %>%
  mutate(wui_bg = case_when(p_nw_hu20 < 50 ~ 1,
                            TRUE ~ 0),
         lint_hu = case_when(pc_hu_den10_20 >= 5 ~ 1,
                             TRUE ~ 0),
         hint_hu = case_when(pc_hu_den10_20 >= 10 ~ 1,
                             TRUE ~ 0),
         lint_pop = case_when(pc_pop_den10_20 >= 5 ~ 1,
                              TRUE ~ 0),
         hint_pop = case_when(pc_pop_den10_20 >= 10 ~ 1,
                              TRUE ~ 0)) %>%
  select(bg_id, wui_bg, lint_hu, hint_hu, lint_pop, hint_pop)
```

Now let's do a spatial join to match block groups to our survey respondents.
```{r}
bg_to_c_point <- st_join(places_to_c_point2, wui_bg1)
bg_to_p_point <- st_join(places_to_p_point2, wui_bg1)
bg_to_w_point <- st_join(places_to_w_point2, wui_bg1)
```

Now let's join our current address data to our previous address data, and then
those joined addresses to the work addresses. We want three sf objects:
1. Current addresses
2. Current addresses + previous addresses (only observations with a previous
address)
3. Current addresses + previous addresses + work addresses (only observations
with a work address)

First, we need to rename variables in previous and work addresses to avoid
duplicate variables after the merge
```{r}
c_addresses <- bg_to_c_point

p_addresses <- bg_to_p_point %>%
  rename(p_county_fips = county_fips,
         p_county = county,
         p_pl_fips = pl_fips,
         p_place = place,
         p_BLK20 = BLK20,
         p_wui20 = wui20,
         p_intermix20 = intermix20,
         p_interfac20 = interfac20,
         p_wui_new20 = wui_new20,
         p_bg_id = bg_id,
         p_wui_bg = wui_bg,
         p_lint_hu = lint_hu,
         p_hint_hu = hint_hu,
         p_lint_pop = lint_pop,
         p_hint_pop = hint_pop,
         p_high_den20 = high_den20,
         p_med_den20 = med_den20,
         p_low_den20 = low_den20
  )

w_addresses <- bg_to_w_point %>%
  rename(w_county_fips = county_fips,
         w_county = county,
         w_pl_fips = pl_fips,
         w_place = place,
         w_BLK20 = BLK20,
         w_wui20 = wui20,
         w_intermix20 = intermix20,
         w_interfac20 = interfac20,
         w_wui_new20 = wui_new20,
         w_bg_id = bg_id,
         w_wui_bg = wui_bg,
         w_lint_hu = lint_hu,
         w_hint_hu = hint_hu,
         w_lint_pop = lint_pop,
         w_hint_pop = hint_pop,
         w_high_den20 = high_den20,
         w_med_den20 = med_den20,
         w_low_den20 = low_den20
  )

```

Now let's join current addresses to previous addresses and current
addresses to work addresses, as well as all addresses to the same sf object
```{r}
cp_addresses <- inner_join(p_addresses, as.data.frame(c_addresses), by = 'ParentGlobalID')
cw_addresses <- inner_join(w_addresses, as.data.frame(c_addresses), by = 'ParentGlobalID')
cpw_addresses <- full_join(c_addresses, as.data.frame(p_addresses), by = 'ParentGlobalID')
cpw_addresses <- full_join(cpw_addresses, as.data.frame(w_addresses, by = 'ParentGlobalID'))
```

# Aggregate to block, block group, and place

So we have data for survey respondents at the point level. Now we want to
aggregate that data up to the block, block group and place scales so that we 
can do choropleth mapping.

Aggregate to block, block group, and places. For now we'll just do current
addresses.

```{r}
point_to_blk <- c_addresses %>%
  group_by(BLK20) %>%
  summarize(n = n()) %>%
  st_drop_geometry()

point_to_bg <- c_addresses %>%
  group_by(bg_id) %>%
  summarize(n = n()) %>%
  st_drop_geometry()

point_to_pl <- c_addresses %>%
  filter(c_inside_cnty == "Yes") %>%
  group_by(place) %>%
  summarize(n = n()) %>%
  st_drop_geometry()
```

Group identical WUI blocks together so that we don't have multiple matches to a
single block when we join (many WUI blocks are partial).
```{r}
wui_blk_consol <- wui_status20 %>%
  group_by(BLK20) %>%
  summarize(blk_frags = n())
```


```{r}
point_to_blk1 <- inner_join(wui_blk_consol, point_to_blk, by = "BLK20")
point_to_bg1 <- inner_join(wui_bg1, point_to_bg, by = "bg_id")
point_to_pl1 <- inner_join(places1, point_to_pl, by = "place")
```

Now let's save our points and each of our polygon files as geojson.
```{r}
st_write(c_addresses, "data/shapes/current_addresses_08-20-24.geojson")
st_write(cp_addresses, "data/shapes/previous_addresses_08-20-24.geojson")
st_write(cw_addresses, "data/shapes/work_addresses_08-20-24.geojson")
st_write(point_to_blk1, "data/shapes/survey_blk_08-20-24.geojson")
st_write(point_to_bg1, "data/shapes/survey_bg_08-20-24.geojson")
st_write(point_to_pl1, "data/shapes/survey_pl_08-20-24.geojson")
```

Let's also take each of our address files, select variables to keep, drop the
geometry, and save them as CSV. These should go back into the main survey 
dataset being worked on.
```{r}
c_addresses1 <- c_addresses %>%
  select(ParentGlobalID, c_partial, county, c_cnty_dist, c_inside_cnty, place, c_pl_dist, 
         c_pl_res, wui20, intermix20, interfac20, wui_new20, wui_bg, lint_hu,
         hint_hu, lint_pop, hint_pop, high_den20, med_den20, low_den20) %>%
  st_drop_geometry()

cp_addresses1 <- cp_addresses %>%
  select(ParentGlobalID, p_partial, p_county, p_cnty_dist, p_inside_cnty, p_place, p_pl_dist, 
         p_pl_res, p_wui20, p_intermix20, p_interfac20, p_wui_new20, p_wui_bg, p_lint_hu,
         p_hint_hu, p_lint_pop, p_hint_pop, p_high_den20, p_med_den20, p_low_den20) %>%
  st_drop_geometry()

cw_addresses1 <- cw_addresses %>%
  select(ParentGlobalID, w_partial, w_county, w_cnty_dist, w_inside_cnty, w_place, w_pl_dist, 
         pl_work, w_wui20, w_intermix20, w_interfac20, w_wui_new20, w_wui_bg, w_lint_hu,
         w_hint_hu, w_lint_pop, w_hint_pop, w_high_den20, w_med_den20, w_low_den20) %>%
  st_drop_geometry()

cpw_addresses1 <- cpw_addresses %>%
  select(ParentGlobalID, c_partial, county, c_cnty_dist, c_inside_cnty, place, c_pl_dist, 
         c_pl_res, wui20, intermix20, interfac20, wui_new20, wui_bg, lint_hu,
         high_den20, med_den20, low_den20, hint_hu, lint_pop, hint_pop, 
         p_partial, p_county, p_cnty_dist, p_inside_cnty, p_place, p_pl_dist, 
         p_pl_res, p_wui20, p_intermix20, p_interfac20, p_wui_new20, p_wui_bg, 
         p_lint_hu, p_hint_hu, p_lint_pop, p_hint_pop, w_partial, w_county, 
         w_cnty_dist, w_inside_cnty, w_place, w_pl_dist, p_high_den20, 
         p_med_den20, p_low_den20, pl_work, w_wui20, w_intermix20, w_interfac20,
         w_wui_new20, w_wui_bg, w_lint_hu, w_hint_hu, w_lint_pop, w_hint_pop, 
         w_high_den20, w_med_den20, w_low_den20) %>%
  st_drop_geometry()
```


Save each address file as CSV (these need work though)
```{r}
#write_csv(c_addresses1, "data/survey/curr_add_wui.csv")
#write_csv(cp_addresses1, "data/survey/prev_add_wui.csv")
#write_csv(cw_addresses1, "data/survey/work_add_wui.csv")
write_csv(cpw_addresses1, "data/survey/all_address_wui.csv")
```


Check out from statistics at various geographies using group_by.
```{r}
c_addresses1 %>%
  group_by(c_pl_res) %>%
  summarize(n = n())
```
